---
title: "Visualization: Bar Charts"
author: Zach del Rosario
date: 2020-05-13
output: github_document
time: 30
reading: 30
---

*Purpose*: *Bar charts* are a key tool for EDA. In this exercise, we'll learn
how to construct a variety of different bar charts, as well as when---and when
*not*---to use various charts.

*Reading*: [Bar Charts](https://rstudio.cloud/learn/primers/3.2)
*Topics*: (All topics)
*Reading Time*: ~30 minutes

```{r setup, include=FALSE}
# knitr options
knitr::opts_chunk$set(echo = TRUE)
```

```{r library}
library(tidyverse)

```

__q1__ In the reading, you learned the relation between `geom_bar()` and
`geom_col()`. Use that knowledge to convert the following `geom_bar()` plot into
the same visual using `geom_col()`.

```{r q1-task}
mpg %>%
  ggplot(aes(trans)) +
  geom_bar()

#my changes 
mpg %>% 
  count(trans) %>% 
  ggplot() +
  geom_col(mapping = aes(x = trans, y = n))
```

The reading mentioned that when using `geom_col()` our x-y data should be
`1-to-1`. This next exercise will probe what happens when our data are not
`1-to-1`, and yet we use a `geom_col()`. Note that a
[one-to-one](https://en.wikipedia.org/wiki/Injective_function) function is one
where each input leads to a single output. For the `mpg` dataset, we can see
that the pairs `cty, hwy` clearly don't have this one-to-one property:

```{r mpg-not-1-to-1}
## NOTE: Run this chunk for an illustration
mpg %>% filter(cty == 20)
```

__q2__ The following code attempts to visualize `cty, hwy` from `mpg` using
`geom_col()`. There's something fishy about the `hwy` values; what's wrong here?

*Hint*: Try changing the `position` parameter for `geom_col()`.

```{r q2-task}
mpg %>% 
  ggplot() + 
  geom_col(mapping = aes(x = cty, y = hwy))

mpg %>%
  ggplot() +
  geom_col(mapping = aes(x = cty, y = hwy, fill = cty), position = "dodge")

mpg %>% 
  ggplot() + 
  geom_col(mapping = aes(x = cty, y = hwy, fill = hwy), position = "stack")
```

**Observations**:
- Even without changing position, the original chart looks like `hwy` speeds go close to 700 mph, which is not reasonable!  I suspect it's multiple values of `hwy` being added on top of each other.
- Adding `position = "dodge"` shows more reasonable values of `hwy`, going up close to 50 mph.  I added color, too, to help me see where the `cty` values were the same, although since `cty` is a continuous variable it didn't help as much as I'd hoped--the gradient is hard to interpret.  Still, at least the values of `hwy` are now visible.
- Changing back to `position = "stack"` and `fill`ing by `hwy` gave me a clear image that, when stacked, it is indeed adding those multiple `hwy` speeds together when they have the same `cty` speed.  So those 600 mph columns are really just multiple cases of <50 mph values being stacked on top!

A more standard way to visualize this kind of data is a *scatterplot*, which
we'll study later. For now, here's an example of a more effective way to
visualize `cty` vs `hwy`:

```{r q2-compare}
## NOTE: Run this chunk for an illustration
mpg %>%
  ggplot(aes(cty, hwy)) +
  geom_point()
```

__q3__ The following are two different visualizations of the `mpg` dataset.
Document your observations between the `v1` and `v2` visuals. Then, determine
which---`v1` or `v2`---enabled you to make more observations. What was the
difference between the two visuals?

```{r q4-v1}
## TODO: Run this code without changing, describe your observations on the data
mpg %>%
  ggplot(aes(class, fill = class)) +
  geom_bar()
```

**Observations**:
- Here I can see how many cars in the `mpg` dataset belong to each `class` of car.  There are more `suv`s (>60) than any other class, and fewer `2seater`s (~ 4 - 5?) than any other class.

```{r q4-v2}
## TODO: Run this code without changing, describe your observations on the data
mpg %>%
  ggplot(aes(class, fill = drv)) +
  geom_bar()
```

**Observations**:
- This now allows me to see everything I observed in the first graph, *plus* the ability to distinguish how many of each class have each kind of `drv`.
- For example, the `suv`s in `mpg` are mostly (like, 80% ish mostly) 4-wheel drive, and the rest are rear-wheel drive -- there aren't any in `mpg` that are front-wheel drive.
- `2seater`s in `mpg` are only rear-wheel drive, and `mpg`'s `pickup`s are only 4-wheel drive.

**Compare `v1` and `v2`**:

- Which visualization---`v1` or `v2`---enabled you to make more observations?
- `v2` allowed me to make more observations.
- What was the difference between `v1` and `v2`?
- `v2` added a `fill` based on `drv`, which allowed me to see a stacked bar chart of the `drv` values for each `class` of car in `mpg`.  `v1` on the other hand used the `fill` aesthetic on `class`, which was already distinguished by `x`.

__q4__ The following code has a bug; it does not do what its author intended.
Identify and fix the bug. What does the resulting graph tell you about the
relation between `manufacturer` and `class` of cars in this dataset?

*Note*: I use a `theme()` call to rotate the x-axis labels. We'll learn how to
do this in a future exercise.

```{r q4-task}
mpg %>%
  ggplot(aes(x = manufacturer, fill = class), position = "dodge") +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust = 0))

# my version
mpg %>%
  ggplot() +
  geom_bar(aes(x = manufacturer, fill = class), position = "dodge") +
  theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust = 0))
```

**Observations**
- The bug was that it wasn't following the `"dodge"` position, which apparently can only be set in the `geom`, not in the `ggplot()` function.
- When moved so that `dodge` was respected, it shows that most `manufacturer`s make a limited number of `class`es of cars that are included in `mpg`.  `jeep`, `land rover`, `lincoln`, and `mercury` only have `SUV`s in this dataset, while `honda` only has `subcompact`s and `pontiac` only has `midsize`.

__q5__ The following graph is hard to read. What other form of faceting would
make the visual more convenient to read? Modify the code below with your
suggested improvement.

```{r q5-task}
mpg %>%
  ggplot(aes(x = cyl)) +
  geom_bar() +
  facet_grid(~ manufacturer)

# my version
mpg %>%
  ggplot(aes(x = cyl)) +
  geom_bar() +
  facet_wrap(~ manufacturer)
```

<!-- include-exit-ticket -->
# Exit Ticket
<!-- -------------------------------------------------- -->

Once you have completed this exercise, make sure to fill out the **exit ticket survey**, [linked here](https://docs.google.com/forms/d/e/1FAIpQLSeuq2LFIwWcm05e8-JU84A3irdEL7JkXhMq5Xtoalib36LFHw/viewform?usp=pp_url&entry.693978880=e-vis01-bar-charts-assignment.Rmd).
